### 2011-02-24 mc/plot-interval.Rnw
Final edits before submission to Econometrica

doxyplot <- function(dplot, ...) {
  print(xyplot(I(1 - reject) ~ I(ntest/n) | interaction(nlabel, altlabel, normlabel, sep = ", "),
               data = dplot,
               ylab = "Coverage", xlab = "P/T",
               panel = function(x,y,...) {
                 panel.xyplot(x,y,...,type = "l")
                 panel.lines(c(0,2/3),c(.9,.9), col = "gray")
               },
               layout = c(4,4),
               par.strip.text = list(cex = .65),
               index.cond = list(c(13,15,16,14,9,11,12,10,5,7,8,6,1,3,4,2))))
}
@

<<testerror1, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'test.error.1', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)

doxyplot(d) #Empirical coverage of one-sided interval
@

<<testerror2, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'test.error.2', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)
doxyplot(d) #Empirical coverage of two-sided interval
@

<<generror1, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'gen.error.1', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)
doxyplot(d) #Empirical coverage of one-sided interval
@

<<generror2, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'gen.error.2', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)
doxyplot(d) #Empirical coverage of two-sided interval
@

>>>>

doxyplot <- function(dplot, ...) {
  print(xyplot(I(1 - reject) ~ I(ntest/n) | interaction(nlabel, altlabel, normlabel, sep = ", "),
               data = dplot,
               ylab = "Coverage", xlab = "P/T",
               panel = function(x,y,...) {
                 panel.lines(c(0,2/3),c(.9,.9), col = "gray")
                 panel.xyplot(x,y,...,type = "l")
               },
               layout = c(4,4),
               par.strip.text = list(cex = .65),
               index.cond = list(c(13,15,16,14,9,11,12,10,5,7,8,6,1,3,4,2))))
}
@

Check empirical coverage of $E_{R} \bar{D}_{1}$ for the one-sided
naive out-of-sample t-test-based interval.
<<testerror1, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'test.error.1', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)

doxyplot(d)
@

Check empirical coverage of $E_{R} \bar{D}_{1}$ for the two-sided
naive out-of-sample t-test-based interval.
<<testerror2, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'test.error.2', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)
doxyplot(d)
@

Check empirical coverage of $E_{T} \bar{D}_{2}$ for the one-sided
naive out-of-sample t-test-based interval.
<<generror1, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'gen.error.1', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)
doxyplot(d) #Empirical coverage of one-sided interval
@

Check empirical coverage of $E_{T} \bar{D}_{2}$ for the two-sided
naive out-of-sample t-test-based interval.
<<generror2, fig=TRUE, height = 3.5, width=6>>=
d <- dbGetQuery(dbc, sprintf(dstring, 'gen.error.2', 500))
d$nlabel <- sprintf("T=%d", d$n)
d$normlabel <- sprintf("c=%d", d$norm)
doxyplot(d) #Empirical coverage of two-sided interval
@

### 2011-02-24 paper.tex
Final edits before submission to Econometrica

\documentclass[11pt]{article}
%\linespread{1.2}\selectfont
\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[round,comma]{natbib}
\usepackage{setspace,amsmath,amssymb,amsthm,ifthen,subfig,url,multicol}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\thepage}\chead{}\rhead{}
\lfoot{}\cfoot{}\rfoot{}
\renewcommand\headrulewidth{0pt}
\onehalfspacing
\usepackage[top=1.222222in,left=0.9444444in,right=1.888889in,bottom=2.4444444in]{geometry}
\usepackage{nath}

@@

\newcommand{\oosA}{\bar{D}_a}
\newcommand{\oosB}{\bar{D}_b}

@@

\frenchspacing \title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Helle Bunzel, Marjorie Flavin,
    Nir Jaimovich, Lutz Kilian, Michael McCracken, Seth Pruitt, Ross
    Starr, Yixiao Sun, Allan Timmermann, participants at the 2010
    Midwest Economics Association Annual Meetings, the 2010 Joint
    Statistical Meetings, and in many seminars at \textsc{ucsd}, and
    espeically Graham Elliott for their valuable suggestions, feedback
    and advice in writing this paper.  I would also like to thank Amit
    Goyal for providing computer code and data for his 2008 RFS paper
    with Ivo Welch \citep{GoW:08}.} \\ Iowa State University,
  Economics Department}
\date{October 24, 2011}

@@

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Under this limit theory, the naive Diebold-Mariano-West
  out-of-sample test can test hypotheses about a key quantity for
  evaluating forecasting models---a time series analogue to the
  generalization error---as long as the out-of-sample period is small
  relative to the total sample size.  Moreover, tests that are
  designed to reject if the larger model is true, such as the usual
  in-sample Wald and \textsc{lm} tests and also Clark and McCracken's
  (2001, 2005a), \citepos{Mcc:07} and Clark and West's (2006, 2007)
  out-of-sample statistics, will choose the larger model too often
  when the smaller model is more accurate.

\noindent \textsc{jel} Classification: C12, C22, C52, C53

\noindent Keywords: Generalization Error, Forecasting, Model
Selection, t-test, Dimension Asymptotics
\end{abstract}

@@

\begin{proof}[Proof of Theorem~\ref{res:ftest}]
  To clean up the notation, we'll write $\theta_1$ as $\theta$ and
  $q_{K_2-K_1,T-K_2}(1-\alpha)$ as $q_\alpha$.  Let
  $(\hat{\theta}, \hat{\psi})$ denote the \ols\ coefficient estimator
  using the full model and let 
  \[\hat\tau^2 = \frac{1}{T-K_2}\sum_{t=0}^{T-1} (y_{t+1} - x_{1t}'\hat{\theta} -
  x_{2t}'\hat{\psi})^2. \] Since the regressors are orthogonal,
  $\hat\theta$ also is the \ols\ estimator of $\theta$ using the
  restricted model.  The difference between the two models'
  generalization error is
\[
E_T \oosB = \sum_{j=K_1+1}^{K_2} [\psi_j^2 - (\hat{\psi}_j - \psi_j)^2]
\]
For any $\hat{\theta}$, $E_T \oosB$ equals zero when
$\hat{\psi}$ lies on the surface of the hypersphere centered at
$\psi$ with radius $|\psi|_2$, and $E_T \oosB$ and is greater
than zero when $\hat{\psi}$ lies outside that sphere.
Also, the rejection region in $\Re^{K_2}$ for is the set of values of $\hat{\psi}$
on or beyond the sphere
\[
\hat{\psi}'\hat{\psi} = \hat{\tau}^2 q_\alpha \frac{K_2-K_1}{T}.
\]

@@

As a result,
\[
\Pr[F \geq q_{\alpha}\wall \mid E_T \oosB \leq 0] = 
\Pr[\hat{\psi}'\hat{\psi} \geq \hat \tau^2 q_\alpha \frac{K_2 - K_1}{T}
\mid |\hat{\psi}-\psi|_2 \geq |\psi|_2] \\
\geq 1 - \E(\Pr\{ \hat{\omega}_i \in [\tilde{\omega}_i - \xi, \tilde{\omega}_i +
\xi] \ \forall i  \mid \hat{\tau}^2,
|\hat{\psi}-\psi|_2 \geq |\psi|_2\} 
\mid |\hat{\psi}-\psi|_2 \geq |\psi|_2)\quad a.s. \\ 
= 1 - \frac12 E ((2\xi / \pi)^{K_2-K_1-1}\mid
|\hat{\psi}-\psi|_2 \geq |\psi|_2)\quad a.s. \\
= 1 - \frac12 E (2\xi/\pi)^{K_2-K_1-1} \quad a.s..
\]
The second equality holds because the $\hat{\omega}_i$s are uniform
and the third holds because $\hat{\tau}$ is independent of
$\hat{\psi}$ under normality.

Now~\eqref{eq:ftestreject1} holds since $\xi \to 0$ almost surely as
$\phi_2 \to \infty.$  Moreover, as
$T\to\infty$, $\hat{\tau}^2\to^p\tau^2$ and $q_{\alpha}\to 1$, so
$\plim \xi < \frac{\pi}{2}$.  Consequently,
$(\frac{2\xi}{\pi})^{K_2-K_1-1}\to^p0$, proving~\eqref{eq:ftestreject2}.
\end{proof}
\bibliographystyle{abbrvnat}
\bibliography{../../texlibrary.bzr/main}

\begin{table}[b]
 \begin{multicols}{2}
  \begin{itemize}
  \item log Dividend Price Ratio
  \item log Earnings Price Ratio
  \item Stock Market Variance
  \item Book to Market Ratio
  \item Net Equity Expansion
  \item Percent Equity Issuing
  \item Treasury Bill
  \item Long Term Yield
  \item Long Term Rate
  \item Default Return Spread
  \item Default Yield Spread
  \item Inflation
  \end{itemize}
 \end{multicols}
 \caption{Variables used in Section \ref{sec:empirics} and by Goyal
   and Welch (2008) to predict the equity premium.  Please see Goyal
   and Welch's original paper
   for a detailed description of each variable.}
 \label{tab:equity}
\end{table}

\begin{table}[!tbp]
 \begin{center}
 \begin{tabular}{lrrrr}\hline\hline
\multicolumn{1}{l}{waldtest}&\multicolumn{1}{c}{Res.Df}&\multicolumn{1}{c}{Df}&\multicolumn{1}{c}{F}&\multicolumn{1}{c}{Pr(\textgreaterF)}\tabularnewline
\hline
1&69&&&\\
2&81&$-$12&3.4&0.00066 \\
\hline
\end{tabular}
\caption{Wald test for the null hypothesis that all of the
  coefficients in Goyal and Welch's (2008) ``kitchen sink'' model are
  zero, except for the intercept, using the Newey-West
  variance-covariance matrix.  The p-value is calculated from the $F$
  distribution.\label{tab:gwinsample}}
\end{center}
\end{table}

@@

\begin{figure}
  \centering
  \large{Simulated Rejection Probability of Clark and West
      (2006, 2007) Test Under $E_T \oosB \leq 0$}
  \includegraphics{mc/plots/size-clarkwest}
  \caption{Simulated rejection probabilities for Clark and West's
    (2006, 2007) \protect\textsc{oos} test statistic given $\E_T \oosB
    \leq 0$ with nominal size 10\%.  Values greater than 10\% indicate
    that the test rejects the benchmark model too often.  The solid
    horizontal line indicates the nominal rejection probability.  See
    Section~\ref{sec:simulation-design} for a discussion of the
    simulation design.}
   \label{fig:clarkwest}
\end{figure}

@@

% LocalWords: LM clark meese diebold chao corradi mccracken giacomini lm jel iR
% analyses inoue datasets huber efron anova akritas
% calhoun iR berbee th de nondifferentiable jong
% mixingales mcleish AIC Berbee's merlevede Heyde's ib
% mixingaleR TP eq dx dy tex tR tS tT iT sR sw tw vw
% reestimate davidson over-reject anatolyev Anatolyev's iS izx
% coefpart normalpart Jong's jt oos underrejects cltvar Scwarz JEL
% hansen lm McCracken's jel RSQLite elemstatlearning rproject rsqlite
% LocalWords:  McCracken's kilian anatolyev rossi REStud JAppliedEconometrics
% LocalWords:  goyal welch rfs reputational calhoun Berbee's mixingales tR tS
% LocalWords:  Heteroskedasticity Autocorrelation tT iT th berbee iS de dx ixz
% LocalWords:  DMW mcleish sR eq ftestreject lightgray veclen RSQLite zeileis
% LocalWords:  heteroskedasticity hothorn RNews JStatSoftware harrell campbell
% LocalWords:  thompson bivariate Welch's multicollinearity Newey newey tw vw
% LocalWords:  merlevede cltvar dR Scwarz davidson indices lrrrr waldtest Df jt
% LocalWords:  dmw McCracken MSE OLS mse inoue huber elemstatlearning efron sw
% LocalWords:  anova akritas jong normalpart coefpart Jong's rproject rsqlite
% LocalWords:  unrestrictive datasets Heyde's ib oos

>>>>

\documentclass[11pt]{article}
%\linespread{1.2}\selectfont
\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{fit}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[round,comma]{natbib}
\usepackage{setspace,amsmath,amssymb,amsthm,ifthen,subfig,url,multicol}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\thepage}\chead{}\rhead{}
\lfoot{}\cfoot{}\rfoot{}
\renewcommand\headrulewidth{0pt}
\onehalfspacing
\usepackage[margin=1in]{geometry}
\usepackage{nath}

@@

\newcommand{\oosA}{\bar{D}_{R}}
\newcommand{\oosB}{\bar{D}_{T}}

@@

\frenchspacing \title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Helle Bunzel, Marjorie Flavin,
    Nir Jaimovich, Lutz Kilian, Michael McCracken, Seth Pruitt, Ross
    Starr, Yixiao Sun, Allan Timmermann, participants at the 2010
    Midwest Economics Association Annual Meetings, the 2010
    International Symposium on Forecasting, the 2010 Joint
    Statistical Meetings, and in many seminars at \textsc{ucsd}, and
    espeically Graham Elliott for their valuable suggestions, feedback
    and advice in writing this paper.  I would also like to thank Amit
    Goyal for providing computer code and data for his 2008 RFS paper
    with Ivo Welch \citep{GoW:08}.} \\ Iowa State University,
  Economics Department}
\date{October 24, 2011}

@@

\begin{abstract}\thispagestyle{empty}\noindent
  This paper uses dimension asymptotics to study why overfit linear
  regression models should be compared out-of-sample; we let the
  number of predictors used by the larger model increase with the
  number of observations so that their ratio remains uniformly
  positive.  Under this limit theory, the naive Diebold-Mariano-West
  out-of-sample test can test hypotheses about a key quantity for
  evaluating forecasting models---a time series analogue to the
  generalization error---as long as the out-of-sample period is small
  relative to the total sample size.  Moreover, tests that are
  designed to reject if the larger model is true, such as the usual
  in-sample Wald and \textsc{lm} tests and also Clark and McCracken's
  (2001, 2005a), \citepos{Mcc:07}, and Clark and West's (2006, 2007)
  out-of-sample statistics, will choose the larger model too often
  when the smaller model is more accurate.

\noindent \textsc{jel} Classification: C01, C12, C22, C52, C53

\noindent Keywords: Generalization Error, Forecasting, Model
Selection, t-test, Dimension Asymptotics
\end{abstract}
\newpage

@@

\begin{proof}[Proof of Theorem~\ref{res:ftest}]
  To clean up the notation, we'll write $\theta_1$ as $\theta$ and
  $q_{K_2-K_1,T-K_2}(1-\alpha)$ as $q_\alpha$.  Let
  $(\hat{\theta}, \hat{\psi})$ denote the \ols\ coefficient estimator
  using the full model and let 
  \[\hat\tau^2 = \frac{1}{T-K_2}\sum_{t=0}^{T-1} (y_{t+1} - x_{1t}'\hat{\theta} -
  x_{2t}'\hat{\psi})^2. \] Since the regressors are orthogonal,
  $\hat\theta$ also is the \ols\ estimator of $\theta$ using the
  restricted model.  The difference between the two models'
  generalization error is
\[
E_T \oosB = \sum_{j=K_1+1}^{K_2} [\psi_j^2 - (\hat{\psi}_j - \psi_j)^2].
\]
For any $\hat{\theta}$, $E_T \oosB$ equals zero when $\hat{\psi}$ lies
on the surface of the hypersphere centered at $\psi$ with radius
$|\psi|_2$, and $E_T \oosB$ and is greater than zero when $\hat{\psi}$
lies outside that sphere.  Also, the rejection region in $\Re^{K_2}$
for the F-test is the set of values of $\hat{\psi}$ on or beyond the
sphere
\[
\hat{\psi}'\hat{\psi} = \hat{\tau}^2 q_\alpha \frac{K_2-K_1}{T}.
\]

@@

As a result,
\[
\Pr[F \geq q_{\alpha}\wall \mid E_T \oosB \leq 0] = 
\Pr[\hat{\psi}'\hat{\psi} \geq \hat \tau^2 q_\alpha \frac{K_2 - K_1}{T}
\mid |\hat{\psi}-\psi|_2 \geq |\psi|_2] \\
\geq 1 - \E(\;\Pr[ \hat{\omega}_i \in [\tilde{\omega}_i - \xi, \tilde{\omega}_i +
\xi] \ \forall i  \mid \wall \hat{\tau}^2,
|\hat{\psi}-\psi|_2 \geq |\psi|_2] \\ 
\mid |\hat{\psi}-\psi|_2 \geq |\psi|_2)\quad a.s. \return \\ 
= 1 - \frac12 E ((2\xi / \pi)^{K_2-K_1-1}\mid
|\hat{\psi}-\psi|_2 \geq |\psi|_2)\quad a.s. \\
= 1 - \frac12 E (2\xi/\pi)^{K_2-K_1-1} \quad a.s..
\]
The second equality holds because the $\hat{\omega}_i$s are uniform
and the third holds because $\hat{\tau}$ is independent of
$\hat{\psi}$ under normality.

Now~\eqref{eq:ftestreject1} holds since $\xi \to 0$ almost surely as
$\phi_2 \to \infty.$  Moreover, as
$T\to\infty$, $\hat{\tau}^2\to^p\tau^2$ and $q_{\alpha}\to 1$, so
$\plim \xi < \frac{\pi}{2}$.  Consequently,
$(\frac{2\xi}{\pi})^{K_2-K_1-1}\to^p0$, proving~\eqref{eq:ftestreject2}.
\end{proof}
\bibliographystyle{abbrvnat}
\bibliography{AllRefs}

\begin{table}[h]
 \begin{multicols}{2}
  \begin{itemize}
  \item log Dividend Price Ratio
  \item log Earnings Price Ratio
  \item Stock Market Variance
  \item Book to Market Ratio
  \item Net Equity Expansion
  \item Percent Equity Issuing
  \item Treasury Bill
  \item Long Term Yield
  \item Long Term Rate
  \item Default Return Spread
  \item Default Yield Spread
  \item Inflation
  \end{itemize}
 \end{multicols}
 \caption{Variables used in Section \ref{sec:empirics} and by Goyal
   and Welch (2008) to predict the equity premium.  Please see Goyal
   and Welch's original paper
   for a detailed description of each variable.}
 \label{tab:equity}
\end{table}

\begin{table}[b]
 \begin{center}
 \begin{tabular}{lrrrr}\hline\hline
\multicolumn{1}{l}{waldtest}&\multicolumn{1}{c}{Res.Df}&\multicolumn{1}{c}{Df}&\multicolumn{1}{c}{F}&\multicolumn{1}{c}{Pr(\textgreaterF)}\tabularnewline
\hline
1&69&&&\\
2&81&$-$12&3.4&0.00066 \\
\hline
\end{tabular}
\caption{Wald test for the null hypothesis that all of the
  coefficients in Goyal and Welch's (2008) ``kitchen sink'' model are
  zero, except for the intercept, using the Newey-West
  variance-covariance matrix.  The p-value is calculated from the $F$
  distribution.\label{tab:gwinsample}}
\end{center}
\end{table}

@@

\begin{figure}
  \centering
  \large{Simulated Rejection Probability of Clark and West
      (2006, 2007) Test \\ Under $E_T \oosB \leq 0$}
  \includegraphics{mc/plots/size-clarkwest}
  \caption{Simulated rejection probabilities for Clark and West's
    (2006, 2007) \protect\textsc{oos} test statistic given $\E_T \oosB
    \leq 0$ with nominal size 10\%.  Values greater than 10\% indicate
    that the test rejects the benchmark model too often.  The solid
    horizontal line indicates the nominal rejection probability.  See
    Section~\ref{sec:simulation-design} for a discussion of the
    simulation design.}
   \label{fig:clarkwest}
\end{figure}

@@

% LocalWords: LM clark meese diebold chao corradi mccracken giacomini lm jel iR
% analyses inoue datasets huber efron anova akritas
% calhoun iR berbee th de nondifferentiable jong
% mixingales mcleish AIC Berbee's merlevede Heyde's ib
% mixingaleR TP eq dx dy tex tR tS tT iT sR sw tw vw
% reestimate davidson over-reject anatolyev Anatolyev's iS izx
% coefpart normalpart Jong's jt oos underrejects cltvar Scwarz JEL
% hansen lm McCracken's jel RSQLite elemstatlearning rproject rsqlite
% LocalWords:  McCracken's kilian anatolyev rossi REStud JAppliedEconometrics
% LocalWords:  goyal welch rfs reputational calhoun Berbee's mixingales tR tS
% LocalWords:  Heteroskedasticity Autocorrelation tT iT th berbee iS de dx ixz
% LocalWords:  DMW mcleish sR eq ftestreject lightgray veclen RSQLite zeileis
% LocalWords:  heteroskedasticity hothorn RNews JStatSoftware harrell campbell
% LocalWords:  thompson bivariate Welch's multicollinearity Newey newey tw vw
% LocalWords:  merlevede cltvar dR Scwarz davidson indices lrrrr waldtest Df jt
% LocalWords:  dmw McCracken MSE OLS mse inoue huber elemstatlearning efron sw
% LocalWords:  anova akritas jong normalpart coefpart Jong's rproject rsqlite
% LocalWords:  unrestrictive datasets Heyde's ib oos Mcc MeR StW InK DiM CCS
% LocalWords:  ClM CoS ClW GiW GiR GoW Cla HTF Efr BoB AkA AkP DeD Mcl Dej Rde
% LocalWords:  Sar Zeh Zei Har CaT NeW MeP HaH AllRefs
